---
title: "SpatialData_and_Mapping"
output:
  pdf_document: default
  html_document: default
---
---
title: "SpatialData_and_Mapping"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{R - Install Packages, message=F}
packages<-c("sf", "raster", "rgdal", "rgeos", "sp", "tidyverse", "rmarkdown","RColorBrewer", "mapview")

lapply(packages,FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)}})

# clean environment
rm(packages)
```

# Working off of Local desktop GIS workspace using Github repository
```{r - Set Working Directory}
# Make sure working directory is automatically set to your local github folder
getwd()

```

Vector data consists of points, lines, and/or polygons with spatial coordinates
# Load Vector Data
```{r - Load Vector Data, message=F}

# Read in dataset of leafy spurge presence points in Fremont Co, WY
SpurgeFull <- readOGR(dsn = ".", layer = "SpurgeFull")
study_area <- readOGR(dsn = ".", layer = "StudyArea")
FremontMRoads <- readOGR(dsn = ".", layer = "FremontMainRoads")
FremontWater <- readOGR(dsn = ".", layer = "WaterwaysFremont")
FremontCo <- readOGR(dsn = ".", layer = "FremontCo")

# Check to make sure data imported properly
head(SpurgeFull)
head(study_area)
head(FremontMRoads)
head(FremontWater)
head(FremontCo)

table(is.na(SpurgeFull$Latitude))# must be FALSE here! No TRUES. 
table(is.na(SpurgeFull$Longitude))# must be FALSE here! No TRUES. 

# Plot data set. May take a bit
plot(SpurgeFull)

# Not that informative, add some spatial context with Fremont Co border and a plot again
```

# break for powerpoint chunk on projection #


# What is the projection that your data are in? Figure this out and specify below. Specify at least 2 different fuctions you can use to check this.
** ** 

``` {r - Projection Systems}
# First, check coordinate system and projection, otherwise these will not align when plotting
proj4string(SpurgeFull)
proj4string(study_area)
proj4string(FremontMRoads)
proj4string(FremontWater)
proj4string(FremontCo)
```

# What happens if your projection systems don't match? How do the two plots below differ?
** ** 

``` {r - Projection mismatch, message=F}
# To see the difference a projection makes
proj <- "+proj=tmerc +lat_0=40.5 +lon_0=-108.75 +k=0.9999375 +x_0=600000 +y_0=0 +datum=NAD83 +units=us-ft +no_defs +ellps=GRS80 +towgs84=0,0,0"
FremontCo <- spTransform(FremontCo, CRS(proj))
par(mfrow = c(1, 2))
plot(FremontCo)

proj <- "+proj=utm +zone=13 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
FremontCo <- spTransform(FremontCo, CRS(proj))
plot(FremontCo)

dev.off()
```


#Change the projection systems so that all data has the same CRS.
```{r - Correct projection, message=F}
proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
SpurgeFull <- spTransform(SpurgeFull, CRS(proj))
study_area <- spTransform(study_area, CRS(proj))
FremontMRoads <- spTransform(FremontMRoads, CRS(proj))
FremontWater <- spTransform(FremontWater, CRS(proj))
FremontCo <- spTransform(FremontCo, CRS(proj))

# make sure all projections line up
proj4string(SpurgeFull)
proj4string(study_area)
proj4string(FremontMRoads)
proj4string(FremontWater)
proj4string(FremontCo)
# perfect!
```

#Can you think of other ways to make sure projection systems are the same? List 1 other way.
** **
``` {r - Projections}

#try some new ways here!

```


```{r - Plotting}
# Plot datasets to make sure they line up. May take a bit
plot(SpurgeFull)
plot(FremontCo, add = T, border = "red")
plot(FremontWater, add = T, border = "blue")
plot(FremontMRoads, add = T, border = "grey")

# You might notice some issues!
```

# What is wrong with the map right now?
** ** 

```{r - Changing extent of map, message=F}

# First lets clip the data to only look at Lander/Riverton area
FremontMRoads <- FremontMRoads[study_area,]
SpurgeFull <- SpurgeFull[study_area,]
FremontWater <- FremontWater[study_area,]

# Check to make sure it worked
plot(SpurgeFull)
plot(study_area, add = T, border = "red")
plot(FremontWater, add = T, border = "blue")
plot(FremontMRoads, add = T, border = "grey")
```

# Why do the roads and rivers still extend outside the study area polygon?
** ** 

```{r - Cleaning up plots}
# Lets make a cleaner version of the intersections between the study area and our water and roads
FremontWater_sub <- crop(FremontWater, study_area)
FremontMRoads_sub <- crop(FremontMRoads, study_area)

plot(SpurgeFull_sub)
plot(study_area, add = T, border = "red")
plot(FremontWater_sub, add = T, border = "blue")
plot(FremontMRoads_sub, add = T, border = "grey")
```

# What has changed in the data? 
# hint:look at road length.
** ** 

```{r - Spatial maps in ggplot, message=F}

# Have to fortify to plot with ggplot
study_areaF <- fortify(study_area)
SpurgeFullF <- fortify(SpurgeFull)
FremontWaterF <- fortify(FremontWater_sub)
FremontCoF <- fortify(FremontCo)
FremontMRoadsF <- fortify(FremontMRoads_sub)

# Plot spurge and Fremont country to continue to get oriented
ggplot() +
  geom_polygon(data = FremontCoF, aes(x=long, y=lat), fill = NA, color = "black") +
  geom_polygon(data = SpurgeFullF, aes(x=long, y=lat, group = group), fill = "NA", color ="green")+
  theme_minimal()+
  labs(x="",y="")+
  theme(panel.grid=element_blank(), axis.text=element_blank())

# Plot data sets with new clipped extent (just our study area)
ggplot() +
  geom_polygon(data = study_areaF, aes(x=long, y=lat), fill = NA, color = "black") +
  geom_polygon(data = FremontWaterF, aes(x=long, y=lat, group = group), fill = "blue", color = "blue" ) +    
  geom_polygon(data = SpurgeFullF, aes(x=long, y=lat, group = group), fill = "yellow", color = "yellow" ) +    
  geom_line(data = FremontMRoadsF, aes(x=long, y=lat, group = group), color = "black" ) +
  theme_minimal()+
  labs(x="", y="")+
  theme(panel.grid = element_blank(), axis.text = element_blank())
```


# Raster data is made up of a grid of pixels filled with values. These can be continuous variables like elevation, thematic values like landcover, or reflectance values like remote sensing imagery. 
```{r - Load raster data}
# For this project, we have one continuous topographic raster and one thematic landcover raster.

# Import elevation
Elevation <- raster("elev_output.tif")

# Import GAP landcover data
GAPCover <- raster("GAP_output.tif")
# Alright, all data is in!

# What to we need to do to make sure things will line up?
proj4string(Elevation)
proj4string(GAPCover)

proj
Elevation <- projectRaster(Elevation, crs = proj)
GAPCover <- projectRaster(GAPCover, crs = proj)

proj4string(Elevation)
proj4string(GAPCover)
# all looks great!

```

# Crop/Mask Raster Data
```{r - Raster crop ,message=F}

# we are going to clip our rasters to match the clipped extent of our other data.
# We have two rasters for you to play with, Elevation, and GAP Cover. 
Elevation_sub <- crop(Elevation, study_area)
GAPCover_sub <- crop(GAPCover, study_area)
```

#What do we need to look at in a Raster other than projection?
** **

```{r - Rasters continued, message=F}

res(Elevation)
res(GAPCover)
# Great! They match

# a few ways to change the raster cell size, here is a simple one for our purposes
Elevation_sub1 <- disaggregate(Elevation_sub, fact = 5, method = "bilinear")

res(Elevation_sub)
res(Elevation_sub1)
# cool... but what does this mean?

par(mfrow = c(1, 2))
plot(Elevation_sub)
plot(Elevation_sub1)
# Looks much better! 

rm(Elevation_sub1)
dev.off()
```

# Why not have a very fine resolution for all rasters all the time?
** **

# Take a look at the raster data again, why do we have to mask it?
** **
  
```{r - Plot Raster Data}
par(mfrow = c(1, 2))
plot(Elevation)
plot(Elevation_sub)
plot(GAPCover)
plot(GAPCover_sub)
```

```{r - Plot all data types together in ggplot}
#Lets start with plotting GAPCover.

#We  need to convert GAP landcover data from a raster to points to use in ggplot
map.p <- rasterToPoints(GAPCover_sub)
df <- data.frame(map.p)

#Make appropriate column headings
colnames(df)<-c("Longitude", "Latitude", "GAPCover")
df<-fortify(df)

#GAPCover map - you'll have to give this a minute or two to plot!
ggplot(df, aes(y=Latitude, x=Longitude)) +
  geom_raster(aes(fill=GAPCover))+
  theme_minimal()+
  theme(panel.grid = element_blank(), axis.text = element_blank())+
  labs(x="", y="")+
  scale_fill_gradientn(colours=brewer.pal(9, "RdPu"))+
  geom_polygon(data = FremontWaterF, aes(x=long, y=lat, group = group), color = "blue")+
  geom_line(data = FremontMRoadsF, aes(x=long, y=lat, group = group), color = "black")+
  geom_polygon(data = SpurgeFullF, aes(x=long, y=lat, group = group), color ="chartreuse")

```

# How to best share maps with colaborators?
```{r - Sharing Maps, message=F}

share_map <- mapview(study_area, fill = FALSE, lwd = 2.2) + mapview(SpurgeFull_sub, map.types = mapviewGetOption("OpenStreetMap")) + mapview(Elevation_sub) + mapview(FremontMRoads_sub) + mapview(FremontCo, fill = FALSE) + mapview(FremontWater_sub)
share_map

####  To make a HTML map to share
#### DO NOT RUN THIS CODE CHUNK #### It will take a long time. It is only an example to work off.

# setwd() # make this to a folder shared with collaborators like google drive or Box
# mapshot(share_map,  url = paste0(getwd(), "/Fremont_County_Spurge.html")) #change map iteration each time unless you want to overwrite. Can provide a good history of maps with changing data

```


# Breakout room task 1
```{r}
# Make a new map (any type) using a combination of the available layers for all of Fremont county AND use this new layer. Detail the steps you need to take in order to make this map and code it out. Try to decrease the cell size to make a smoother looking map with any raster layers.

FremontAgencies <- readOGR(dsn = ".", layer = "FremontAgencies")

```


# Breakout room task 2. 
# Try making a map of elevation and water that fits inside of all Fremont County. 
```{r - Lets try changing the extent to which our map is bound ,message=F}
rm(list = ls())

#Re-read in your spatial data
GAPCover <- raster("GAP_output.tif")
Elevation <- raster("elev_output.tif")
SpurgeFull <- readOGR(dsn = ".", layer = "SpurgeFull")
FremontMRoads <- readOGR(dsn = ".", layer = "FremontMainRoads")
FremontWater <- readOGR(dsn = ".", layer = "WaterwaysFremont")
FremontCo <- readOGR(dsn = ".", layer = "FremontCo")

  #1st mask it first to Fremont Co!
 
# -----Your code here!----- # 

  # Next you will need to convert the raster to points- try it!

# -----Your code here!----- # 

  # Last, plot it all w/ggplot! (you can use the GAPCover code for reference!)

# -----Your code here!----- # 

```
