---
title: "SpatialData_and_Mapping"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
---
title: "SpatialData_and_Mapping"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Courtney
```{R - Install Packages, message=F}
#Install and Library for individual packages, if needed
#install.packages("sf")
#install.packages("raster")
#install.packages("rgdal")
#install.packages("rgeos")
#install.packages("sp")
#install.packages("tidyverse")
#install.packages("rmarkdown")
#install.packages("RColorBrewer")
#install.packages("mapview")
#install.packages("readr")
#library(sf)
#library(raster)
#library(rgdal)
#library(rgeos)
#library(sp)
#library(tidyverse)
#library(rmarkdown)
#library(RColorBrewer)
#library(mapview)
#library(readr)

packages<-c("sf", "raster", "rgdal", "rgeos", "sp", "tidyverse", "rmarkdown","RColorBrewer", "mapview", "readr")

lapply(packages,FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)}})

# clean environment
rm(packages)
```

# Working off of Local desktop GIS workspace using Github repository
# Courtney
```{r - Set Working Directory}
# Make sure working directory is automatically set to your local github folder
getwd()

```

Vector data consists of points, lines, and/or polygons with spatial coordinates
# Load Vector Data
# Courtney
```{r - Load Vector Data, message=F}

# Read in dataset of leafy spurge presence points in Fremont Co, WY
SpurgePoint <- readOGR(dsn = ".", layer = "SpurgePoint")

# Polygon data, polygons, polylines, etc.
study_area <- readOGR(dsn = ".", layer = "StudyArea")
FremontMRoads <- readOGR(dsn = ".", layer = "FremontMainRoads")
FremontWater <- readOGR(dsn = ".", layer = "WaterwaysFremont")
FremontCo <- readOGR(dsn = ".", layer = "FremontCo")

# Check to make sure data imported properly
head(SpurgePoint)
head(study_area)
head(FremontMRoads)
head(FremontWater)
head(FremontCo)

table(is.na(SpurgePoint$Latitude))# must be FALSE here! No TRUES. 
table(is.na(SpurgePoint$Longitude))# must be FALSE here! No TRUES. 

# Plot data set. May take a bit
plot(SpurgePoint)

# Not that informative as a standalone map, we will add some spatial context after we read in the next data
```

# Now, read in raster data...
# Raster data is made up of a grid of pixels filled with values. These can be continuous variables like elevation, thematic values like landcover, or reflectance values like remote sensing imagery. 
# Courtney
```{r - Load raster data}
# For this project, we have one continuous topographic raster and one thematic landcover raster.

# Import elevation
Elevation <- raster("elev_output.tif")

# Import GAP landcover data
GAPCover <- raster("GAP_output.tif")

# Alright, all data is in!
```


# break for powerpoint chunk on projection #


# What is the projection that your data are in? Figure this out and specify below. Specify at least 2 different fuctions you can use to check this.
** ** 

# Nathan
``` {r - Projection Systems}
# First, check coordinate system and projection, otherwise these will not align when plotting
# proj4string(SpurgeFull)
proj4string(SpurgePoint)
proj4string(study_area)
proj4string(FremontMRoads)
proj4string(FremontWater)
proj4string(FremontCo)
```

# What happens if your projection systems don't match? How do the two plots below differ?
** ** 

# Nathan
``` {r - Projection mismatch, message=F}
# To see the difference a projection makes
proj <- "+proj=tmerc +lat_0=40.5 +lon_0=-108.75 +k=0.9999375 +x_0=600000 +y_0=0 +datum=NAD83 +units=us-ft +no_defs +ellps=GRS80 +towgs84=0,0,0"
FremontCo <- spTransform(FremontCo, CRS(proj))
par(mfrow = c(1, 2))
plot(FremontCo)

proj <- "+proj=utm +zone=13 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
FremontCo <- spTransform(FremontCo, CRS(proj))
plot(FremontCo)

dev.off()
```


#Change the projection systems so that all data has the same CRS.
# Nathan
```{r - Correct projection, message=F}
proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
SpurgePoint <- spTransform(SpurgePoint, CRS(proj))
study_area <- spTransform(study_area, CRS(proj))
FremontMRoads <- spTransform(FremontMRoads, CRS(proj))
FremontWater <- spTransform(FremontWater, CRS(proj))
FremontCo <- spTransform(FremontCo, CRS(proj))

# make sure all projections line up
proj4string(SpurgePoint)
proj4string(study_area)
proj4string(FremontMRoads)
proj4string(FremontWater)
proj4string(FremontCo)
# perfect!
```

# Now, do the same projection changes for the raster data, Elevation and GAPCover
# Nathan
```{r - Raster Projections}
# What to we need to do to make sure things will line up?
proj4string(Elevation)
proj4string(GAPCover)

proj
Elevation <- projectRaster(Elevation, crs = proj)
GAPCover <- projectRaster(GAPCover, crs = proj)

proj4string(Elevation)
proj4string(GAPCover)
# all looks great!
```


#Can you think of other ways to make sure projection systems are the same? List 1 other way.
** **

# Nathan
``` {r - Projections}

#try some new ways here!

```

#Plot time
# Chloe
```{r - Plotting, message=F}
# Plot data sets to make sure they line up. May take a bit
plot(SpurgePoint)
plot(FremontCo, add = T, border = "red")
plot(FremontWater, add = T, border = "blue")
plot(FremontMRoads, add = T, border = "grey")

# You might notice some issues!
```

# What is wrong with the map right now?
** ** 

# Spatial Data Processing - Vector Data
# Chloe
```{r - Changing extent of map, message=F}

# First lets clip the data to only look at Lander/Riverton area
FremontMRoads <- FremontMRoads[study_area,]
SpurgePoint <- SpurgePoint[study_area,]
FremontWater <- FremontWater[study_area,]

# Check to make sure it worked
plot(SpurgePoint)
plot(study_area, add = T, border = "red")
plot(FremontWater, add = T, border = "blue")
plot(FremontMRoads, add = T, border = "grey")
```

# Why do the roads and rivers still extend outside the study area polygon?
** ** 

# Chloe
```{r - Cleaning up plots}
# Lets make a cleaner version of the intersections between the study area and our water and roads
FremontWater_sub <- crop(FremontWater, study_area)
FremontMRoads_sub <- crop(FremontMRoads, study_area)
SpurgePoint_sub <- crop(SpurgePoint, study_area)

plot(SpurgePoint_sub)
plot(study_area, add = T, border = "red")
plot(FremontWater_sub, add = T, border = "blue")
plot(FremontMRoads_sub, add = T, border = "grey")
```

# What has changed in the data? 
# hint:look at road length.
** ** 

# Sarah
```{r - Spatial maps in ggplot, message=F}

# Have to fortify to plot with ggplot
study_areaF <- fortify(study_area)
# Make SpurgePoints a dataframe
SpurgePointdf <- as.data.frame(SpurgePoint)
SpurgePointF <- fortify(SpurgePointdf)
FremontWaterF <- fortify(FremontWater_sub)
FremontCoF <- fortify(FremontCo)
FremontMRoadsF <- fortify(FremontMRoads_sub)

# Plot spurge and Fremont country to continue to get oriented
ggplot() +
  geom_polygon(data = FremontCoF, aes(x=long, y=lat), fill = NA, color = "black") +
  geom_point(data = SpurgePointF, aes(x=Longitude, y=Latitude), fill = "NA", color = "green") +
  theme_minimal()+
  labs(x="",y="")+
  theme(panel.grid=element_blank(), axis.text=element_blank())

# Plot data sets with new clipped extent (just our study area)
ggplot() +
  geom_polygon(data = study_areaF, aes(x=long, y=lat), fill = NA, color = "black") +
  geom_polygon(data = FremontWaterF, aes(x=long, y=lat, group = group), fill = "blue", color = "blue" ) + 
  geom_point(data = SpurgePointF, aes(x=Longitude, y=Latitude), fill = "yellow", color = "yellow" ) +
  geom_line(data = FremontMRoadsF, aes(x=long, y=lat, group = group), color = "black" ) +
  theme_minimal()+
  labs(x="", y="")+
  theme(panel.grid = element_blank(), axis.text = element_blank())
```

# Processing of Spatial Data - Raster
# Crop/Mask Raster Data and Spurge Points
# Chloe
```{r - Raster crop ,message=F}

# we are going to clip our rasters to match the clipped extent of our other data.
# We have two rasters for you to play with, Elevation, and GAP Cover. 
Elevation_sub <- crop(Elevation, study_area)
GAPCover_sub <- crop(GAPCover, study_area)
```

#What do we need to look at in a Raster other than projection?
** **

# Chloe
```{r - Rasters continued, message=F}

res(Elevation)
# res(Slope)
res(GAPCover)
# Great! They match

# a few ways to change the raster cell size, here is a simple one for our purposes
Elevation_sub1 <- disaggregate(Elevation_sub, fact = 5, method = "bilinear")

res(Elevation_sub)
res(Elevation_sub1)
# cool... but what does this mean?

par(mfrow = c(1, 2))
plot(Elevation_sub)
plot(Elevation_sub1)
# Looks much better! 

rm(Elevation_sub1)
dev.off()
```

# Why not have a very fine resolution for all rasters all the time?
** **

# Take a look at the raster data again, why do we have to mask it?
** **
# Chloe
```{r - Plot Raster Data}
par(mfrow = c(1, 2))
plot(Elevation)
plot(Elevation_sub)
plot(GAPCover)
plot(GAPCover_sub)
# Let's dig into what this GAPCover raster is representing...
```

# Read in Look Up Table of Sorts for GAPCover Data
# Chloe
```{r - Raster Look Up Table for GAP, message = F}
plot(GAPCover)
# Where are these values in the legend coming from? And what do they mean?
# Let's check it out by reading in a data table that corresponds to GAPCover values
GAP_LUT <- read.csv("GAP_LU.txt", header = TRUE, sep = ",", dec = ".")
# Explore this data table

```

# What column(s) are we pulling values from when we plot GAPCover
** **

# Describe differences between the three CN_LEVELS 
** **

# Sarah
```{r - Plot all data types together in ggplot, Elevation}
#Lets start with plotting Elevation.

#We  need to convert GAP landcover data from a raster to points to use in ggplot
map.p <- rasterToPoints(Elevation_sub)
df <- data.frame(map.p)

#Make appropriate column headings
colnames(df)<-c("Longitude", "Latitude", "Elevation")
df<-fortify(df)

#Elevation map - you'll have to give this a minute or two to plot!
ggplot(df, aes(y=Latitude, x=Longitude)) +
  geom_raster(aes(fill=Elevation))+
  theme_minimal()+
  theme(panel.grid = element_blank(), axis.text = element_blank())+
  labs(x="", y="")+
  scale_fill_gradientn(colours=brewer.pal(9, "RdPu"))+
  geom_polygon(data = FremontWaterF, aes(x=long, y=lat, group = group), color = "blue")+
  geom_line(data = FremontMRoadsF, aes(x=long, y=lat, group = group), color = "black")+
  #geom_polygon(data = SpurgeFullF, aes(x=long, y=lat, group = group), color ="chartreuse")
  geom_point(data = SpurgePointF, aes(Longitude, y=Latitude, color = "chartreuse"))

```


# How to best share maps with collaborators?
# Nathan
```{r - Sharing Maps, message=F}

# share_map <- mapview(study_area, fill = FALSE, lwd = 2.2) + mapview(SpurgeFull_sub, map.types = mapviewGetOption("OpenStreetMap")) + mapview(Elevation_sub) + mapview(FremontMRoads_sub) + mapview(FremontCo, fill = FALSE) + mapview(FremontWater_sub)
# share_map

share_map <- mapview(study_area, fill = FALSE, lwd = 2.2) + mapview(SpurgePoint_sub, map.types = mapviewGetOption("OpenStreetMap")) + mapview(Elevation_sub) + mapview(FremontMRoads_sub) + mapview(FremontCo, fill = FALSE) + mapview(FremontWater_sub)
share_map

####  To make a HTML map to share
#### DO NOT RUN THIS CODE CHUNK #### It will take a long time. It is only an example to work off.

# setwd() # make this to a folder shared with collaborators like google drive or Box
# mapshot(share_map,  url = paste0(getwd(), "/Fremont_County_Spurge.html")) #change map iteration each time unless you want to overwrite. Can provide a good history of maps with changing data

```


# Breakout room task  
# Try making a map of elevation and water that fits inside of all Fremont County. 
# All
```{r - Lets try changing the extent to which our map is bound ,message=F}

# Make a new map (any type) using a combination of the available layers for all of Fremont county AND use this new layer, slope_output.tif. 
Slope <- raster("slope_output.tif")
#Detail the steps you need to take in order to make this map and code it out. Try to decrease the cell size to make a smoother looking map with any raster layers.

rm(list = ls())

#Re-read in your spatial data
GAPCover <- raster("GAP_output.tif")
Elevation <- raster("elev_output.tif")
SpurgePoints <- readOGR(dsn = ".", layer = "SpurgePoints")
FremontMRoads <- readOGR(dsn = ".", layer = "FremontMainRoads")
FremontWater <- readOGR(dsn = ".", layer = "WaterwaysFremont")
FremontCo <- readOGR(dsn = ".", layer = "FremontCo")

  #1st mask it first to Fremont Co!
 
# -----Your code here!----- # 

  # Next you will need to convert the raster to points- try it!

# -----Your code here!----- # 

  # Last, plot it all w/ggplot! (you can use the GAPCover code for reference!)

# -----Your code here!----- # 

```
